---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# robsurvey

The package **robsurvey** provides several functions to compute robust survey statistics. The package supports the computations of robust means, totals, and ratios. Available methods are Huber M-estimators, trimming, and winsorization. **robsurvey** complements the famous [survey package](https://cran.r-project.org/web/packages/survey/index.html).

## Overview

The following functions are provided in **robsurvey**:

* `weighted.median()` 
* `weighted.quantile()`
* `weighted.mad()`
* `weighted.total()`
* `weighted.mean()`
* `msvymean()`
* `msvytotal()`
* `tsvymean()`
* `tsvytotal()`
* `wsvymean()`
* `wsvytotal()`

## Installation

You can install robsurvey from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("martinSter/robsurvey")
```

## Example

In the following example, we showcase a typical use of the package **robsurvey**. The data we use are from the package **survey** and describe the student performance in California schools. We will show different ways of how to compute a robust mean value for the Academic Performance Index (API) in 2000. The variable is denoted as `api00`. The following code chunk simply loads the data and defines the survey design (based on the **survey** package).

```{r example}
## load packages
library(robsurvey)
library(survey)

# load the api dataset
data(api)

# define survey design
dstrat <- svydesign(id = ~1, strata = ~stype, weights = ~pw, 
                    data = apistrat, fpc = ~fpc)
```

In the following code chunk we first compute the robust Horvitz-Thompson M-estimator for the mean. In addition, we compute the trimmed and winsorized (robust) Horvitz-Thompson mean. Note how the estimates and the corresponding standard errors vary.

```{r}
# compute the robust Horvitz-Thompson M-estimator of the mean
msvymean(~api00, dstrat, k = 2)

# compute the robust trimmed Horvitz-Thompson mean
tsvymean(~api00, dstrat, k = 2)

# compute the robust winsorized Horvitz-Thompson mean
wsvymean(~api00, dstrat, k = 2)
```

It is also possible to use `msvymean()` in combination with `svyby()` from the **survey** package. The variable `stype` denotes the school level: elementary, middle, and high school.

```{r}
# Domain estimates
svyby(~api00, by = ~stype, design = dstrat, msvymean, k = 1.34)
```

